Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-25 overleaf/tech_report/results.tex

\subsection{Experiment Setup}\label{sec:experiment_setup}
We tested our framework with several benchmarks. For the multiparty computation (MPC), we restricted our evaluation to 2 party computation (2PC) setting because it requires fewer computing resources. We stress that there is no such inherent restriction in our framework. We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Network (LAN) and Wide Area Network (WAN). In the LAN setting, we use {\tt c6525-25g} machines for both parties. These machines are equipped with 16-core AMD 7302P 3.0GHz processors and 128GB of RAM. The connection between these machines had 10Gbps bandwidth and sub-millisecond latency. %This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks. 
In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second. The {\tt c220g1} machine is equipped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM. We measured the connection bandwidth to be 560Mbps and average round trip time (RTT) to be 38ms. %At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.

We run all experiments 5 times and report average values of various metrics. Note that standard deviation in all observations was at most 4.5\% of the mean value. %,  therefore the accuracy of the results is not effected by the relatively fewer runs.

\subsection{Benchmarks}\label{sec:benchmarks_description}

We used benchmarks from ABY and OPA~\cite{Demmler:2015,Ishaq:2019} (Biometric Matching, Histogram, PSI), HyCC~\cite{Buscher:2018b} (Database Join, Database Variance, Inner Product, Cryptonets Max Pooling, MNIST ReLU, k-means Iteration) and Parsynt~\cite{Farzan:2021} (Convex Hull, Minimal Points, Count 102, Count 10, Longest 102, Max distance between symbols). 
We did not use Gcd~\cite{Ishaq:2019} and Modular Exponentiation from~\cite{Demmler:2015} as they do not benefit from vectorization according to our model. 
We ran as many routines from~\cite{Buscher:2018} and~\cite{Farzan:2021} as possible. Gaussian decomposition cannot be expressed 
in our syntax and the compiler errors on Cryptonets Convolution at the time of writing, although we expect to fix the issue soon. 
\cite{Anon_TR} describes what the benchmarks do.

To compare we run {\em both} versions with the maximal value of $N$ for which the non-vectorized version succeeds (typically {\tt N=128}). We also run {\em vec} with larger values, 
typically {\tt N=4096}. Again, due to space constraints \cite{Anon_TR} gives the exact values.
% In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
%In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value for {\em both} and we fixed this value for these experiments. The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons).
% however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).

@@

\begin{comment}
\begin{enumerate}
    \item {\em Biometric Matching:} Server has a database S of {\tt N} records, each record's dimension is {\tt D}. Client submits a query C, client and server compute the closest record to C in an MPC. We use {\tt N=128} for {\em both} and {\tt N=4096} for {\em vec}. {\tt D} is fixed at 4.

@@

    \item {\em Private Set Intersection (PSI)} Alice holds set S1 with size {\tt SA}, Bob holds set S2 with size {\tt SB}, together they compute intersection of their sets. We use {\tt SA=SB=128} for {\em both} and {\tt SA=SB=1024} for {\em vec}.
\end{enumerate}
\end{comment}

@@

{\small
   \begin{table*}[htbp]
       \centering
       \caption{Vectorized vs Non-Vectorized Comparison, times in seconds (in LAN setting where applicable), Communication in MiB, Numbers in 1000s, values rounded to nearest integer, benchmark names ending in {\em V} are vectorized.}
       \label{table:metrics}
       \rowcolors{5}{}{gray!10}
       \tabcolsep3pt
       \begin{tabular}{lrrrrrrrrrrrr}
           \toprule
           {} & \multicolumn{6}{c}{\textbf{GMW}} & \multicolumn{6}{c}{\textbf{BMR}} \\
            \cmidrule(r){2-7} \cmidrule(r){8-13} \\
            {\bf Benchmark} & Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. &  Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. \\
          \midrule
            Biometric Matching & 146 & 16 & 1,784 & 119 & 1,413 & 140 & 89 & 263 & 1,595 & 139 & 2,716 & 312\\
            Biometric Matching (V) & 12 & 4 & 34 & 2 & 28 & 14 & 2 & 13 & 30 & 4 & 61 & 130\\
            \midrule
            Convex Hull & 48 & 6 & 551 & 40 & 516 & 51 & 28 & 72 & 494 & 39 & 695 & 80\\
            Convex Hull (V) & 0 & 1 & 2 & 0 & 1 & 4 & 0 & 2 & 1 & 1 & 2 & 32\\
            \midrule
            Count 102 & 79 & 6 & 418 & 35 & 525 & 52 & 15 & 62 & 269 & 33 & 785 & 92\\
            Count 102 (V) & 71 & 5 & 316 & 24 & 332 & 34 & 11 & 30 & 167 & 16 & 304 & 59\\
            \midrule
            Count 10s & 79 & 6 & 419 & 35 & 525 & 52 & 14 & 62 & 270 & 33 & 785 & 92\\
            Count 10s (V) & 71 & 4 & 316 & 24 & 332 & 34 & 11 & 29 & 167 & 16 & 304 & 59\\
            \midrule
            Cryptonets (Max Pooling) & 50 & 11 & 688 & 46 & 554 & 55 & 36 & 89 & 608 & 51 & 898 & 110\\
            Cryptonets (Max Pooling) (V) & 1 & 1 & 7 & 1 & 2 & 5 & 2 & 4 & 7 & 2 & 12 & 49\\
            \midrule
            Database Join & 70 & 8 & 433 & 48 & 790 & 80 & 19 & 229 & 458 & 119 & 3,518 & 427\\
            Database Join (V) & 54 & 6 & 320 & 35 & 575 & 61 & 16 & 112 & 320 & 57 & 1,457 & 285\\
            \midrule
            Database Variance & 166 & 18 & 2,009 & 135 & 1,639 & 163 & 95 & 269 & 1,708 & 145 & 2,795 & 320\\
            Database Variance (V) & 37 & 6 & 321 & 24 & 334 & 43 & 10 & 30 & 170 & 13 & 178 & 141\\
            \midrule
            Histogram & 94 & 10 & 862 & 68 & 979 & 97 & 27 & 94 & 491 & 51 & 1,132 & 135\\
            Histogram (V) & 33 & 5 & 166 & 16 & 164 & 23 & 7 & 17 & 92 & 13 & 154 & 68\\
            \midrule
            Inner Product & 127 & 15 & 1,675 & 108 & 1,308 & 130 & 83 & 250 & 1,526 & 134 & 2,623 & 301\\
            Inner Product (V) & 16 & 5 & 158 & 12 & 165 & 25 & 6 & 18 & 83 & 7 & 86 & 127\\
            \midrule
            k-means & 108 & 12 & 1,333 & 88 & 1,090 & 108 & 63 & 185 & 1,141 & 99 & 1,958 & 225\\
            k-means (V) & 6 & 3 & 47 & 4 & 43 & 12 & 2 & 11 & 32 & 4 & 54 & 95\\
            \midrule
            Longest 102 & 93 & 7 & 650 & 52 & 713 & 71 & 26 & 93 & 475 & 49 & 1,091 & 128\\
            Longest 102 (V) & 169 & 6 & 544 & 41 & 519 & 53 & 25 & 60 & 369 & 33 & 605 & 95\\
            \midrule
            Max. Dist. b/w Symbols & 71 & 8 & 572 & 43 & 576 & 57 & 24 & 69 & 397 & 38 & 748 & 89\\
            Max. Dist. b/w Symbols (V) & 166 & 7 & 538 & 39 & 512 & 51 & 24 & 57 & 363 & 32 & 589 & 78\\
            \midrule
            Minimal Points & 35 & 5 & 458 & 31 & 369 & 37 & 24 & 46 & 401 & 26 & 347 & 40\\
            Minimal Points (V) & 0 & 1 & 1 & 0 & 1 & 3 & 0 & 1 & 1 & 0 & 1 & 16\\
            \midrule
            MNIST ReLU & 132 & 31 & 1,843 & 126 & 1,483 & 152 & 98 & 247 & 1,630 & 135 & 2,401 & 298\\
            MNIST ReLU (V) & 3 & 3 & 25 & 3 & 9 & 17 & 5 & 11 & 25 & 5 & 33 & 136\\
            \midrule
            Private Set Intersection & 95 & 9 & 558 & 59 & 1,049 & 104 & 22 & 186 & 591 & 96 & 2,639 & 302\\
            Private Set Intersection (V) & 1 & 2 & 1 & 2 & 1 & 8 & 1 & 8 & 2 & 4 & 2 & 122\\
          \bottomrule
       \end{tabular}
       \label{table:metrics}
   \end{table*}
}
\begin{figure*}[htbp]
\centering
%\resizebox{7in}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
\label{fig:graph_all_eval_time}
\end{figure*}

A detailed summary of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}. We show circuit evaluation times in \cref{fig:graph_all_eval_time}. In terms of observed benefits from vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include Convex Hull, Cryptonets Max Pooling, Minimal Points and Private Set Intersection. 
%These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 
2) {\it Medium:} these include Biometric Matching, DB Variance, Histogram, Inner Product, k-means Iteration and MNIST ReLU. These benchmarks have non-parallelizable phases e.g. the summing phase of Inner Product. Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol. 3) {\it Low:} these include the Database Join and the regular expression benchmarks (Count 102, Count 10, Longest 102 and Max distance between symbols). There is less parallelizable computation in these programs, thus the speedup is lower. We see a speedup from 1.1x to 2x in BMR. In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x. However, Longest 102 and Max distance between symbols suffer a slowdown of 0.5x. There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized. We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW. The cause of the increase is not clear --- we conjecture that MOTION performs optimization over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization. We propose a simple heuristic: if the transformation increases circuit depth beyond some threshold (e.g. more than 10\% of the original circuit), we can reject the transformation. Nevertheless, we show these benchmarks here for the sake of completeness and to highlight that vectorization does not always reduce run time. Note that in some settings it may still be desirable to vectorize e.g. in data constrained environments as communication as well as all other metrics, is reduced for all benchmarks. (The extended version presents plots of for all metrics~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).

%We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively. %\ishaq{Ana: Let me know if I should add more substance.} \ana{It's good!}

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
\caption{Biometric Matching Circuit Evaluation Time}
\label{fig:graph_biometric_eval_time}
\end{figure}

\begin{comment}
\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-CommunicationMiB}}
\caption{Biometric Matching Communication Size, x-axis lists database size}
\label{fig:graph_biometic_comm_size}
\end{figure}

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-CircuitGenerationTimesec}}
\caption{Biometric Matching Circuit Generation Time, x-axis lists database size}
\label{fig:graph_biometic_circ_gen_time}
\end{figure}
\end{comment}

We look closely at the Biometric Matching benchmark: circuit evaluation in~\cref{fig:graph_biometric_eval_time}, communication size in~\cite{Anon_TR} and circuit generation time in~\cite{Anon_TR}.
%\cref{fig:graph_biometic_comm_size} and circuit generation time in~\cref{fig:graph_biometic_circ_gen_time}. 
For input size beyond {\tt N=128} the memory usage exceeds available memory and prevents circuit generation. Consequently, non-vectorized bars are missing beyond this threshold. Notice that vectorization improves all metrics. %A database of size {\tt N=128} is smaller than any real world use case, therefore lets use it as baseline for improvement. 
Comparing performance improvement between BMR and GMW, we see more speedup for BMR (23x vs 10x), GMW gets more communication size reduction (10x vs 2.5x) and circuit generation sees a speedup of 35x and 45x for BMR and GMW respectively.

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/comparison-hist-OnlineSetupTimesec}}
\caption{LAN vs. WAN: Circuit Evaluation Time}
\label{fig:graph_comparison_eval_time}
\end{figure}

@@

\begin{comment}
\begin{figure*}[htbp]
\centering
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-CommunicationMiB}}
\caption{Communication Size of Benchmarks}
\label{fig:graph_comm_size}
\end{figure*}

@@

\begin{figure*}[htbp]
\centering
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-SetupTimesec}}
\caption{Setup Time of Benchmarks}
\label{fig:graph_setup_time}
\end{figure*}
\end{comment}

\subsection{Comparison with MOTION-native Inner Product and Discussion}

%We conclude the . 
Finally, we compared our automatically generated routine for Inner Product with the manually SIMD-ified MOTION-native routine in the distribution. 
We were surprised that we were an order of magnitude slower in Boolean GMW as our circuit ran a significantly larger 
number of communication rounds. Upon investigation, it turned out that the vectorized multiplications were essentially the same, 
however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
{\sf result += mult\_unsimdified[i];} while our loop ran auto generated {\sf result[i] = result\_prev[i] + mult\_unsimdified[i];} (as scalar expansion
is an artifact of our vectorization). We rewrote the accumulation manually and that led to the comparable speedup!

MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to
limitations of our implementation. Recall that Phase 3 of Vectorization in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that static analysis 
will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above simpler code. 
However, because AST analysis is difficult, it does not appear to detect the accumulation.
%(unfortunately, we ran out of time and did not experiment further).
%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.

We conjecture that MPC Source, a straight-forward representation suitable for static analysis, will not only enable detection of general associative loops, but also allow 
for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this and other optimizations as future work.

HyCC~\cite{Buscher:2018b} is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$ (as its focus is mixing). %; it does provide a lot of data on mixed protocol executions.
We estimate we are about an order of magnitude slower, which is likely due to the same issue --- the computation of $\mathit{min}$ can be optimized.

>>>>

\subsection{Experiment Setup}\label{sec:experiment_setup}
We tested our framework with several benchmarks. For the multiparty computation (MPC), we restricted our evaluation to 2 party computation (2PC) setting because it requires fewer computing resources. We stress that there is no such inherent restriction in our framework. We use hardware resources provided by CloudLab\cite{DuplyakinATC19} and consider two network settings, namely Local Area Network (LAN) and Wide Area Network (WAN). In the LAN setting, we use {\tt c6525-25g} machines for both parties. These machines are equipped with 16-core AMD 7302P 3.0GHz processors and 128GB of RAM. The connection between these machines had 10Gbps bandwidth and sub-millisecond latency. This setting reflects typical LAN use case considering that 10Gbps LAN is increasingly common in business networks and is now available even in some home networks. In the WAN setting we again used a {\tt c6525-25g} machine (located in Utah, US) for the first party and a {\tt c220g1} machine (located in Wisconsin, US) for the second. The {\tt c220g1} machine is equipped with two Intel E5-2630 8-core 2.40GHz processors and 128GB of RAM. We measured the connection bandwidth between these machines to be 560Mbps and average round trip time (RTT) to be 38ms. At the time of this writing, all major internet providers in the US offer 1Gbps connections to home consumers, therefore this setting reasonably reflects the typical WAN use case.

We run all experiments 5 times and report average values of various metrics. Note that standard deviation --- shown as error bar on top of the bars in the graphs --- in all observations was at most 4.5\% of the mean value, therefore the accuracy of the results is not effected by the relatively fewer runs.

\subsection{Benchmarks}\label{sec:benchmarks_description}
In the following, we say {\em both} for an experiment in which we execute both non-vectorized and vectorized protocols and {\em vec} for the vectorized only experiment. 
In a preliminary experiment we ran {\tt N=2, N=4}, all the way to {\tt N=4096}; the non-vectorized version ran out of memory at the value {\em both} and we fixed this value for these experiments. The vectorized one completed for {\tt N=4096} for nearly all benchmarks (numbers are not shown for space reasons; however, they times are largely consistent, e.g., if {\em both} is $N = 2^k$ and it completes in $X$ seconds for vectorized, then $N = 2^{12}$ completes in $(12-k)X$ seconds).
We used the following benchmarks in our evaluation:

@@

\begin{enumerate}
    \item {\em Biometric Matching:} Server has a database S of {\tt N} records, each record's dimension is {\tt D}. Client submits a query C, client and server compute the closest record to C in an MPC. We use {\tt N=128} for {\em both} and {\tt N=4096} for {\em vec}. {\tt D} is fixed at 4.

@@

    \item {\em Private Set Intersection (PSI)} Alice holds set S1 with size {\tt SA}, Bob holds set S2 with size {\tt SB}, together they compute intersection of their sets. We use {\tt SA=SB=128} for {\em both} and {\tt SA=SB=1024} for {\em vec}.
\end{enumerate}

@@

   \begin{table*}[htbp]
       \centering
       \caption{Vectorized vs Non-Vectorized Comparison, times in seconds (in LAN setting where applicable), Communication in MiB, Numbers in 1000s, values rounded to nearest integer, benchmark names ending in {\em V} are vectorized.}
       \label{table:metrics}
       \rowcolors{5}{}{gray!10}
       \tabcolsep3pt
       \begin{tabular}{lrrrrrrrrrrrr}
           \toprule
           {} & \multicolumn{6}{c}{\textbf{GMW}} & \multicolumn{6}{c}{\textbf{BMR}} \\
            \cmidrule(r){2-7} \cmidrule(r){8-13} \\
            {\bf Benchmark} & Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. &  Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. \\
          \midrule
            Biometric Matching & 146 & 16 & 1,784 & 119 & 1,413 & 140 & 89 & 263 & 1,595 & 139 & 2,716 & 312\\
            Biometric Matching (V) & 12 & 4 & 34 & 2 & 28 & 14 & 2 & 13 & 30 & 4 & 61 & 130\\
            \midrule
            Convex Hull & 48 & 6 & 551 & 40 & 516 & 51 & 28 & 72 & 494 & 39 & 695 & 80\\
            Convex Hull (V) & 0 & 1 & 2 & 0 & 1 & 4 & 0 & 2 & 1 & 1 & 2 & 32\\
            \midrule
            Count 102 & 79 & 6 & 418 & 35 & 525 & 52 & 15 & 62 & 269 & 33 & 785 & 92\\
            Count 102 (V) & 71 & 5 & 316 & 24 & 332 & 34 & 11 & 30 & 167 & 16 & 304 & 59\\
            \midrule
            Count 10s & 79 & 6 & 419 & 35 & 525 & 52 & 14 & 62 & 270 & 33 & 785 & 92\\
            Count 10s (V) & 71 & 4 & 316 & 24 & 332 & 34 & 11 & 29 & 167 & 16 & 304 & 59\\
            \midrule
            Cryptonets (Max Pooling) & 50 & 11 & 688 & 46 & 554 & 55 & 36 & 89 & 608 & 51 & 898 & 110\\
            Cryptonets (Max Pooling) (V) & 1 & 1 & 7 & 1 & 2 & 5 & 2 & 4 & 7 & 2 & 12 & 49\\
            \midrule
            Database Join & 70 & 8 & 433 & 48 & 790 & 80 & 19 & 229 & 458 & 119 & 3,518 & 427\\
            Database Join (V) & 54 & 6 & 320 & 35 & 575 & 61 & 16 & 112 & 320 & 57 & 1,457 & 285\\
            \midrule
            Database Variance & 166 & 18 & 2,009 & 135 & 1,639 & 163 & 95 & 269 & 1,708 & 145 & 2,795 & 320\\
            Database Variance (V) & 37 & 6 & 321 & 24 & 334 & 43 & 10 & 30 & 170 & 13 & 178 & 141\\
            \midrule
            Histogram & 94 & 10 & 862 & 68 & 979 & 97 & 27 & 94 & 491 & 51 & 1,132 & 135\\
            Histogram (V) & 33 & 5 & 166 & 16 & 164 & 23 & 7 & 17 & 92 & 13 & 154 & 68\\
            \midrule
            Inner Product & 127 & 15 & 1,675 & 108 & 1,308 & 130 & 83 & 250 & 1,526 & 134 & 2,623 & 301\\
            Inner Product (V) & 16 & 5 & 158 & 12 & 165 & 25 & 6 & 18 & 83 & 7 & 86 & 127\\
            \midrule
            k-means & 108 & 12 & 1,333 & 88 & 1,090 & 108 & 63 & 185 & 1,141 & 99 & 1,958 & 225\\
            k-means (V) & 6 & 3 & 47 & 4 & 43 & 12 & 2 & 11 & 32 & 4 & 54 & 95\\
            \midrule
            Longest 102 & 93 & 7 & 650 & 52 & 713 & 71 & 26 & 93 & 475 & 49 & 1,091 & 128\\
            Longest 102 (V) & 169 & 6 & 544 & 41 & 519 & 53 & 25 & 60 & 369 & 33 & 605 & 95\\
            \midrule
            Max. Dist. b/w Symbols & 71 & 8 & 572 & 43 & 576 & 57 & 24 & 69 & 397 & 38 & 748 & 89\\
            Max. Dist. b/w Symbols (V) & 166 & 7 & 538 & 39 & 512 & 51 & 24 & 57 & 363 & 32 & 589 & 78\\
            \midrule
            Minimal Points & 35 & 5 & 458 & 31 & 369 & 37 & 24 & 46 & 401 & 26 & 347 & 40\\
            Minimal Points (V) & 0 & 1 & 1 & 0 & 1 & 3 & 0 & 1 & 1 & 0 & 1 & 16\\
            \midrule
            MNIST ReLU & 132 & 31 & 1,843 & 126 & 1,483 & 152 & 98 & 247 & 1,630 & 135 & 2,401 & 298\\
            MNIST ReLU (V) & 3 & 3 & 25 & 3 & 9 & 17 & 5 & 11 & 25 & 5 & 33 & 136\\
            \midrule
            Private Set Intersection & 95 & 9 & 558 & 59 & 1,049 & 104 & 22 & 186 & 591 & 96 & 2,639 & 302\\
            Private Set Intersection (V) & 1 & 2 & 1 & 2 & 1 & 8 & 1 & 8 & 2 & 4 & 2 & 122\\
          \bottomrule
       \end{tabular}
       \label{table:metrics}
   \end{table*}

\begin{figure*}[htbp]
\centering
%\resizebox{7in}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
\label{fig:graph_all_eval_time}
\end{figure*}

A detailed summary of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}. We show circuit evaluation times in \cref{fig:graph_all_eval_time}. In terms of amenability to vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include convex hull, cryptonets max pooling, minimal points and private set intersection. These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 2) {\it Medium:} these include biometric matching, DB Variance, histogram, inner product, k-means iteration and MNIST ReLU. These benchmarks have non-parallelizable phases e.g. the summing phase of inner product and biometric matching. Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol. 3) {\it Low:} these include the Database Join and the regular expression benchmarks (count 102, count 10, longest 102 and max distance between symbols). There is less parallelizable computation in these programs, thus the speedup is lower. We see a speedup from 1.1x to 2x in BMR. In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x. However, longest102 and max distance between symbols suffer a slowdown of 0.5x. There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized. We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW. The cause of the increase is not clear --- we conjecture that MOTION performs optimizations over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization. We propose a simple heuristic although we do leave all the benchmarks in the table: if the transformation increases circuit depth beyond some threshold (e.g. more than 10\% of the original circuit), we can reject the transformation. %Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time. 
Note that in some settings it may still be desirable to vectorize e.g. in data constrained environments. As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).

We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively. %\ishaq{Ana: Let me know if I should add more substance.} \ana{It's good!}

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-OnlineSetupTimesec}}
\caption{Biometric Matching Circuit Evaluation Time, x-axis lists database size}
\label{fig:graph_biometric_eval_time}
\end{figure}

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-CommunicationMiB}}
\caption{Biometric Matching Communication Size, x-axis lists database size}
\label{fig:graph_biometic_comm_size}
\end{figure}

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/lan/biometric-hist-CircuitGenerationTimesec}}
\caption{Biometric Matching Circuit Generation Time, x-axis lists database size}
\label{fig:graph_biometic_circ_gen_time}
\end{figure}

We look closely at the Biometric Matching benchmark: circuit evaluation in~\cref{fig:graph_biometric_eval_time}, communication size in \cref{fig:graph_biometic_comm_size} and circuit generation time in~\cref{fig:graph_biometic_circ_gen_time}. For input size beyond {\tt N=128} the memory usage exceeds available memory and prevents circuit generation. Consequently, non-vectorized bars are missing beyond this threshold. Notice that vectorization improves all metrics. %A database of size {\tt N=128} is smaller than any real world use case, therefore lets use it as baseline for improvement. 
Comparing performance improvement between BMR and GMW, we see more speedup for BMR (23x vs 10x), GMW gets more communication size reduction (10x vs 2.5x) and circuit generation sees a speedup of 35x and 45x for BMR and GMW respectively.

\begin{figure}[htbp]
\centering
\resizebox{0.7\columnwidth}{!}{\input{graphs/comparison-hist-OnlineSetupTimesec}}
\caption{LAN vs. WAN: Circuit Evaluation Time Comparison}
\label{fig:graph_comparison_eval_time}
\end{figure}

@@

\begin{figure*}[htbp]
\centering
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-CommunicationMiB}}
\caption{Communication Size of Benchmarks}
\label{fig:graph_comm_size}
\end{figure*}

@@

\begin{figure*}[htbp]
\centering
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-SetupTimesec}}
\caption{Setup Time of Benchmarks}
\label{fig:graph_setup_time}
\end{figure*}

\subsection{Comparison with MOTION-native Inner Product}

%We conclude the . 
Finally, we compared our automatically generated routine for Inner Product with the manually SIMD-ified MOTION-native routine in the distribution. 
We were surprised that we were an order of magnitude slower in Boolean GMW as our circuit ran a significantly larger 
number of communication rounds. Upon investigation, it turned out that the vectorized multiplications were essentially the same, 
however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i-1] + mult\_unsimdified[i];} recall that the scalar expansion
is an artifact of our vectorization. We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!

MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
MOTION does the standard divide-and-conquer accumulation. The poor performance of our loop is not a limitation of our approach, but it could be a 
limitation of our current implementation. Recall that the next phase of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
will detect the associative accumulation in the scalar expansion code, it is realistic to expect that it will in the above only slightly more complex code. 
It still might lead to confusion in the static analysis as analysis on the AST is difficult (unfortunately, we ran out of time and did not experiment further).
%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.

We conjecture that MPC Source, a straight-forward representation, will not only enable detection of general associative loops, but also allow 
for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this as future work.

The most closely related work, HyCC~\cite{CCS:BDKKS18}, is a mixing compiler that focuses on mixed protocols, while we run vectorization within a single protocol. 
The paper only provides data for Boolean and Yao for a version of Biometric matching on $N=1000$; it does provide a lot of data on mixed protocol executions.
Again, we estimate we are about an order of magnitude slower, which is likely due to the same issue as Inner Product --- the computation of $\mathit{min}$ can be optimized.
